### 2024-11-12 index.qmd
Update index.qmd

---
title: Regional-scale Base-flow Index of Gauged and Ungauged Basins in Drylands
author:
  - name: Caelum Mroczek
    orcid: 0009-0005-6695-0134
    corresponding: true
    email: csm428@nau.edu
    roles:
      - Conceptualization
      - Investigation
      - Methodology
      - Formal analysis
      - Visualization
      - Software
      - Data curation
      - Writing - Original Draft
    affiliations:
      - School of Earth and Sustainability, Northern Arizona University, Flagstaff, AZ, USA

  - name: Abraham Springer
    orcid: 0000-0003-4826-9124
    corresponding: false
    roles:
      - Conceptualization
      - Supervision
      - Project administration
      - Funding acquisition
    affiliations:
      - School of Earth and Sustainability, Northern Arizona University, Flagstaff, AZ, USA

  - name: Benjamin Lucas
    orcid: 0000-0002-2021-3076
    corresponding: false
    roles:
      - Software
      - Supervision
    affiliations:
      - Department of Mathematics and Statistics, Northern Arizona University, Flagstaff, AZ, USA

keywords:
  - Base Flow
  - Base-flow Index
  - Regionalization
  - Machine Learning

abstract: |
  LOREM IPSUM ######
plain-language-summary: |
  LOREM IPSUM ######

key-points:
  - Base flow sustains dryland river ecology, but assessing its groundwater contribution faces data limitations
  - Recursive filtering and machine learning predict BFI and streamflow-associated recharge
  -

date: last-modified
bibliography: references.bib
citation:
  container-title: Water Resources Research
number-sections: true
---

## Introduction {#sec-intro}

Dryland regions, encompassing arid, semi-arid, hyper-arid, and dry sub-humid systems, account for 40% of the Earth’s land surface. These regions are home to approximately 2 billion people globally and constitute the largest terrestrial biome [@iucn-drylands-2019]. Despite supporting diverse ecosystems and human populations, dryland regions face mounting challenges exacerbated by increasing urbanization, expanding agricultural activities, and climate-induced amplification of precipitation patterns [@taylor2013]. This water scarcity is intensifying due to the compounding effects of climate variability and increased groundwater extraction [@taylor2013]. In drylands, groundwater serves as a vital resource for sustaining ecosystems and meeting human needs [@scanlon2006; @yao2018].

Base flow is the sustained portion of streamflow in the absence of runoff that is derived from groundwater discharge [@usgs-glossary-2018]. Base flow is critical to maintaining seasonal low-flow regimes, supporting aquatic ecosystems, and facilitating the transport of nutrients and chemicals. Base-flow contribution to streamflow can be highly variable spatially [@singh2018; @bosch2017; @beck2013], and temporally [@ficklin2016; @tan2020]. Increasing groundwater extraction, changes in land cover/land use, and changes in precipitation patterns due to climate change affect the timing and volumes of base flow [@tan2020; @taylor2013]. Effective management of water quantity and quality depends on understanding seasonal and interannual base flow patterns and long-term changes in base flow behavior.

The Base-Flow Index (BFI) is the ratio of the long-term mean base-flow volume to the long-term total streamflow volume expressed as a percentage. BFI serves as a normalized measure of groundwater contribution interannually or between basins. BFI is determined by hydrograph separation and is influenced by the climate and physiographic characteristics of a catchment [@neff2005; @beck2013; @singh2018]. Between catchments, base flow fluctuates according to changes in the moisture content of the vadose zone, influenced by varying levels of evapotranspiration and aquifer storage dynamics [@bosch2017]. Since BFI calculations rely on instrumented stream records, it remains unknown for ungauged catchments, which encompass most of the earth's land surface [@fekete2007]. Addressing this information gap is integral to approaching a comprehensive understanding of groundwater dynamics globally.

To address the challenge of quantifying base flow in ungauged catchments, numerous studies have applied regression and machine learning methods. [@ahiablame2013] found that using a regression model to estimate annual base flow of ungauged catchments was reasonably easy and accurate. @beck2013 overcame the nonlinearity of basin characteristics and improved results of multivariate analyses by using artificial neural networks to estimate BFI globally. @singh2018 implemented a random forest algorithm to predict long-term BFI for ungauged catchments across New Zealand. The application of machine learning in past studies highlights its versatility and effectiveness in addressing ecohydrologic dynamics.

This study develops a technique for estimating BFI in ungauged basins in Arizona and determines the state’s long-term mean BFI from 1991 to 2020. Due to interannual variability in BFI, a 30-year period of record is used to evaluate the characteristic base-flow dynamics of Arizona. Using a machine learning model trained with the hydrogeologic characteristics of surface water basins, we first determine annual BFIs for each year in the 1991-2020 study period for ungauged river reaches and physiographic regions of Arizona. Long-term mean BFI data for the study period are then estimated. Regional patterns and trends in base flow of instrumented sites are subsequently investigated. These results provide novel analyses of low flow processes in Arizona...######

## Data & Methods {#sec-data-methods}

### Study Area {#sec-study-area}

The state of Arizona, located in the southwest United States, covers a total area of 295,253 km$^2$. Arizona is divided into two primary physiographic provinces: the Colorado Plateau in the northeast, and the Basin-and-Range region in the northwest and south. The Central Highlands is a transition zone consisting of scattered basins separated by the mountainous foothills of the Mogollon Rim. The Colorado Plateau is dominated by high-elevation desert with an average elevation of 1,936 masl (6,352 ft). Average temperatures range from -6$^\circ$C (20$^\circ$F) to 26$^\circ$C (80$^\circ$F) and it averages 580 mm (23 in) of precipitation. The Basin-and-Range region has a semi-arid to arid climate with an average elevation of 490 masl (1600 ft). The average precipitation is 200 mm (8 in) annually with average temperatures ranging from 15$^\circ$C (60$^\circ$F) to 43$^\circ$C (110$^\circ$F).

Arizona's hydrology varies seasonally and spatially between its physiographic regions. In the summer, localized and intense convective precipitation events are driven by the North American Monsoon, while in the winter, orographic precipitation comes from Pacific frontal systems [@eastoe2019mtnblock]. While monsoonal precipitation can account for up to 50% of annual precipitation, evaporation and dry preceding soil properties leads to most precipitation becoming runoff [@sheppard2002]. As such, 94% of streams in Arizona are ephemeral or intermittent [@epa_AZephemeral]. Much of the hydrology of Arizona is snow-melt derived, driven by spring melt from the high-elevation Colorado Plateau winter snowpack. While winter precipitation provides only 30% of annual averages, it provides the bulk of water for natural reservoirs [@sheppard2002].

### Data {#sec-data}

Daily observed streamflow data obtained from the United States Geological Survey (USGS) National Water Information System (NWIS) were used in this study. Streamgages were selected depending on criteria to ensure the applicability of each site. @odonnell2016 found that 8-10 years of calibration data were necessary to account for climate variability in paired watershed studies in the region. Following this criteria, a minimum of 10 years of record was chosen. Any years missing more than 30 days of streamflow were removed from the possible record period. This analysis focuses on natural, streamflow-influencing recharge, excluding streamgages affected by regulation or diversions. These gages were identified through annual reports on water data published by the USGS [@usgs2010]. Finally, streamgages on the Colorado River were excluded from the dataset on the basis that they represent managed flows shared by Colorado River Compact states. Following these parameters, 206 USGS streamgages with acceptable periods of record were used in this study (Figure \###### study-area).

Watersheds across the United States are delineated by the USGS using a hydrologically-defined network. This system delineates the country using hierarchical hydrologic unit codes (HUCs), where each subsequent basin includes the digits of the enclosing basin. Here, 8-digit HUCs (HUC 8s) are used to divide Arizona into 84 sub-basins that are fully or partially in the state (Figure \###### study-area). These HUC 8 sub-basins are analogous to medium-sized river basins and are defined by surface water characteristics. \######

The annual precipitation and temperature data came from the PRISM climate group at Oregon State University at a resolution of 4 km (<https://prism.oregonstate.edu;> [@daly2008]). The PRISM dataset provides valuable insights into regional climate in ungauged regions and has been shown to perform well across the southwestern US [@buban_PRISM]**.** Instead of the water year, PRISM data uses a calendar-year format, which was adopted for consistency in the water balance. Although this may introduce challenges in the annual estimates due to inter-annual snow storage, the use of long-term annual averages is likely to reduce any potential errors [@reitz2017].

Annual evapotranspiration (ET) data came from TerraClimate, a 4-km grid climatalogical data set [@abatzoglou2018]. TerraClimate uses a Penman-Monteith approach to generate a reference potential evapotranspiration (PET). The PET values were calculated assuming a reference grass surface across the landscape with unlimited water. In the drylands of the southwestern US, PET typically exceeds precipitation annually [@zomer2022].

A 30-meter resolution Digital Elevation Model (DEM) of Arizona was used to derive key basin characteristics: basin area, average slope, and the proportion of each basin oriented toward north or south aspects. Various geospatial variables, such as aspect, were disaggregated then averaged to assess the areal percentage of each sub-variable within individual HUC 8 basins. By calculating these percentages, we aimed to get a more comprehensive understanding of landscape composition across space. Land cover from USGS-NLCD (National Land Cover Database), hydrologic soil group from SSURGO (Soil Survey Geographic Database), and underlying geology/karst from USGS were all similarly averaged across the basins. Aggregating variables to align with the HUC 8 boundaries allowed for more precise predictions of BFI by integrating spatial variations within each basin.

### Base-flow separation {#sec-bf_sep}

Directly measuring base flow and BFI presents unique challenges [@eckhardt2008]. The technique chosen to separate base flow has been shown to affect results, and the choice of base-flow separation method is subjective since ‘true’ BFI values are not known [@beck2013]. However, many methods have been developed to estimate these values. These methods include the use of tracers [@gonzales2009], graphical interpolation [@instituteofhydrology1980; @sloto1996], and digital filters [@arnold1995; @eckhardt2005; @lyne1979; @nathan1990]. These techniques have varying levels of applicability depending on the spatial scale, time span, and the scope of the study. Comparisons of various base-flow separation techniques have been made in [@eckhardt2005; @eckhardt2008; @nathan1990] , but this study does not explore the superiority of different methods.

Base flow was calculated using a single-parameter, recursive digital filter technique from [@nathan1990]. This base-flow separation technique is based on a recursive digital filter used in signal analysis that separates high-frequency signals (quickflow) from low-frequency signals (base flow) [@lyne1979]. This technique has been used in [@ARNOLD200021; @santhi2008; @bloomfield2009; @singh2018]. It takes the form of

$$
b = \alpha b_{k-1} + \frac{1-\alpha}{2}(Q_k + Q_{k-1})
$$ {#eq-bf-separate}

where $b$ is base flow, $\alpha$ is the filter parameter, $Q$ is the total streamflow, and $k$ is the time step. A filter parameter $\alpha$ of 0.925 was used as in @nathan1990 and @fuka2014ecohydrology. The filter was run three times (forward, backward, forward) to attenuate the base-flow signal. @eckhardt2023 noted that recursive digital filters lack a physical basis, but as the method is easy to automate, objective, and repeatable, it is appropriate for a regional-scale study.

### Machine Learning

The implementation of machine learning models to predict hydrologic indices has been successful in past studies [@singh2018; @schmidt_ML_2020; @Rozos2021Machine]. In this work, we used the eXtreme Gradient Boosting (XGBoost) algorithm **(chen2016xgboost)** to predict BFI at ungauged locations using catchment characteristics as predictors. The XGBoost algorithm is a decision tree-based ensemble algorithm, which can be adapted for either regression or classification problems. It has grown in popularity in recent years due to its rapid, efficient, and scalable nature, and has proven effective in similar environmental modeling tasks such as streamflow forecasting [@szczepanek2022daily**; @ni2020streamflow**] and classifying land use/land cover data [@Georganos2018Very].

XGBoost uses gradient boosting on decision tree algorithms to leverage the combination of multiple very low variance models into one overall prediction. Gradient boosting works sequentially, with the first tree being learned on the known target values, and each subsequent tree being learned on the errors of the tree prior. The algorithm then applies a weighting to each tree based on this boosting to determine how much the tree will contribute towards the final output. The final prediction is obtained by the input data traversing all $n$ weighted trees in the learned ensemble. The trained XGBoost model is used to predict BFI in ungauged catchments based on geospatial and hydroclimate predictor variables (Table \ref{tab-######}##}##}##}##}##}. Certain predictors were subdivided according to their areal coverage of each basin (e.g. land cover was divided into 16 subdivisions).

|              | Variable                   | Source | Geoprocessing          |
|--------------|----------------------------|--------|------------------------|
| Hydroclimate | Precipitation              | PRISM  | Basin average          |
|              | Mean Temperature           | PRISM  | Basin average          |
|              | Average Evapotranspiration | MODIS  | Basin average          |
| Geospatial   | Elevation                  | DEM    | Basin average          |
|              | Area                       | DEM    | Basin average          |
|              | Slope                      | DEM    | Basin average          |
|              | Aspect                     | DEM    | Percent areal coverage |
|              | Land Cover                 | NLCD   | Percent areal coverage |
|              | Hydrologic Soil Group      | SSURGO | Percent areal coverage |
|              | Geology                    | USGS   | Percent areal coverage |
|              | Karst                      | USGS   | Percent areal coverage |

: **Table 1.** Basin-characteristic variables used as predictors in XGBoost model.

## Results

### BFI of Gauged Catchments

|                              | n    | R2    | MSE    | RMSE  | MAE   | Nash-Sutcliffe | pbias |
|---------|---------|---------|---------|---------|---------|---------|---------|
| Climate - Monsoon Dominated  | 3345 | 0.707 | 0.0166 | 0.129 | 0.077 | 0.694          | -12.6 |
| Climate - Snowmelt Dominated | 6152 | 0.723 | 0.017  | 0.131 | 0.094 | 0.715          | -3.7  |
| PhysRegion - Basin&Range     | 7754 | 0.749 | 0.018  | 0.133 | 0.089 | 0.741          | -5.9  |
| PhysRegion - CO Plateau      | 1743 | 0.806 | 0.013  | 0.115 | 0.081 | 0.801          | -4.4  |
| Climate - Warm-Wet           | 1613 | 0.690 | 0.015  | 0.124 | 0.081 | 0.679          | -9.9  |
| Climate - Warm-Dry           | 3134 | 0.758 | 0.023  | 0.151 | 0.097 | 0.745          | -8.1  |
| Climate - Cool-Wet           | 3134 | 0.710 | 0.013  | 0.115 | 0.085 | 0.708          | -2.5  |
| Climate - Cool-Dry           | 1616 | 0.809 | 0.014  | 0.118 | 0.082 | 0.803          | -5    |
| Slope - High                 | 4398 | 0.753 | 0.014  | 0.120 | 0.084 | 0.746          | -4.4  |
| Slope - Low                  | 5099 | 0.759 | 0.019  | 0.138 | 0.090 | 0.750          | -6.9  |

### BFI of Ungauged Catchments

#### Predictor Importance

#### Model Validation

#### Predicted Long-term Mean BFI

## Conclusion {#sec-conclusion}

## References {.unnumbered}

::: {#refs}
:::