### 2024-12-08 MANUSCRIPT.qmd
Writing
Methods - ML + Feature Selection

Results - Ungauged model validation + importance

---
title: Regional-scale Base-flow Index of Gauged and Ungauged Basins in Drylands
author:
  - name: Caelum Mroczek
    orcid: 0009-0005-6695-0134
    corresponding: true
    email: csm428@nau.edu
    roles:
      - Conceptualization
      - Investigation
      - Methodology
      - Formal analysis
      - Visualization
      - Software
      - Data curation
      - Writing - Original Draft
    affiliations:
      - School of Earth and Sustainability, Northern Arizona University, Flagstaff, AZ, USA

  - name: Abraham Springer
    orcid: 0000-0003-4826-9124
    corresponding: false
    roles:
      - Conceptualization
      - Supervision
      - Project administration
      - Funding acquisition
    affiliations:
      - School of Earth and Sustainability, Northern Arizona University, Flagstaff, AZ, USA

  - name: Benjamin Lucas
    orcid: 0000-0002-2021-3076
    corresponding: false
    roles:
      - Software
      - Supervision
    affiliations:
      - Department of Mathematics and Statistics, Northern Arizona University, Flagstaff, AZ, USA

keywords:
  - Base Flow
  - Base-flow Index
  - Regionalization
  - Machine Learning

abstract: |
  LOREM IPSUM ######
plain-language-summary: |
  LOREM IPSUM ######

key-points:
  - Base flow sustains dryland river ecology, but assessing its groundwater contribution faces data limitations
  - Recursive filtering and machine learning predict BFI 
  -

date: last-modified
bibliography: references.bib
citation:
  container-title: Water Resources Research
number-sections: true
---

## Introduction {#sec-intro}

Dryland regions, encompassing arid, semi-arid, hyper-arid, and dry sub-humid systems, account for 40% of the Earth’s land surface. These regions are home to approximately 2 billion people globally and constitute the largest terrestrial biome [@iucn-drylands-2019]. Despite supporting diverse ecosystems and human populations, dryland regions face mounting challenges exacerbated by increasing urbanization, expanding agricultural activities, and climate-induced amplification of precipitation patterns [@taylor2013]. This water scarcity is intensifying due to the compounding effects of climate variability and increased groundwater extraction [@taylor2013]. In drylands, groundwater serves as a vital resource for sustaining ecosystems and meeting human needs [@scanlon2006; @yao2018].

Base flow is the sustained portion of streamflow in the absence of runoff that is derived from groundwater discharge [@usgs-glossary-2018]. Base flow is critical to maintaining seasonal low-flow regimes, supporting aquatic ecosystems, and facilitating the transport of nutrients and chemicals. Base-flow contribution to streamflow can be highly variable spatially [@singh2018; @bosch2017; @beck2013], and temporally [@ficklin2016; @tan2020]. Increasing groundwater extraction, changes in land cover/land use, and changes in precipitation patterns due to climate change affect the timing and volumes of base flow [@tan2020; @taylor2013]. Effective management of water quantity and quality depends on understanding seasonal and interannual base flow patterns and long-term changes in base flow behavior.

The Base-Flow Index (BFI) is the ratio of the long-term mean base-flow volume to the long-term total streamflow volume expressed as a percentage. BFI serves as a normalized measure of groundwater contribution interannually or between basins. BFI is determined by hydrograph separation and is influenced by the climate and physiographic characteristics of a catchment [@neff2005; @beck2013; @singh2018]. Between catchments, base flow fluctuates according to changes in the moisture content of the vadose zone, influenced by varying levels of evapotranspiration and aquifer storage dynamics [@bosch2017]. Since BFI calculations rely on instrumented stream records, it remains unknown for ungauged catchments, which encompass most of the earth's land surface [@fekete2007]. Addressing this information gap is integral to approaching a comprehensive understanding of groundwater dynamics globally.

To address the challenge of quantifying base flow in ungauged catchments, numerous studies have applied regression and machine learning methods. [@ahiablame2013] found that using a regression model to estimate annual base flow of ungauged catchments was reasonably easy and accurate. @beck2013 overcame the nonlinearity of basin characteristics and improved results of multivariate analyses by using artificial neural networks to estimate BFI globally. @singh2018 implemented a random forest algorithm to predict long-term BFI for ungauged catchments across New Zealand. The application of machine learning in past studies highlights its versatility and effectiveness in addressing ecohydrologic dynamics.

This study develops a technique for estimating BFI in ungauged basins in Arizona and determines the state’s long-term mean BFI. Using a machine learning model trained with the hydrogeologic characteristics of surface water basins, we estimate mean BFI in ungauged basins over a period of record from 1991-2020. Regional patterns and trends in base flow and BFI of instrumented sites are subsequently investigated. These results provide novel analyses of low flow processes in Arizona, furthering our understanding of climate controls on consistent flows in dryland basins.

## Data & Methods {#sec-data-methods}

### Study Area {#sec-study-area}

The state of Arizona, located in the southwest United States, covers a total area of 295,253 km$^2$. Arizona is divided into two primary physiographic provinces: the Colorado Plateau in the northeast, and the Basin-and-Range region in the west and south. The Central Highlands is a transition zone consisting of scattered basins separated by the mountainous foothills of the Mogollon Rim. The Colorado Plateau is dominated by high-elevation desert with an average elevation of 1,936 masl (6,352 ft). Mean temperatures range from -6$^\circ$C (20$^\circ$F) to 26$^\circ$C (80$^\circ$F) and it averages 580 mm (23 in) of precipitation. The Basin-and-Range region has a semi-arid to arid climate with an average elevation of 490 masl (1600 ft). Average temperature ranges from 15$^\circ$C (60$^\circ$F) to 43$^\circ$C (110$^\circ$F) and the region averages 200 mm (8 in) of precipitation annually. ###CITATION###

![Map of Arizona indicating US Geological Survey (USGS) streamgages used in this study. 8-digit HUC subbasin boundaries and physiographic regions shown.](images/StudyArea_20241203.png){#fig-study-area}

Arizona's hydrology varies seasonally and spatially between its physiographic regions. In the summer, localized and intense convective precipitation events are driven by the North American Monsoon, while in the winter, orographic precipitation comes from Pacific frontal systems [@eastoe2019mtnblock]. While monsoonal precipitation can account for up to 50% of annual precipitation, evaporation and dry preceding soil properties leads to most precipitation becoming runoff [@sheppard2002]. As such, 94% of streams in Arizona are ephemeral or intermittent [@epa_AZephemeral]. Much of the hydrology of Arizona is snow-melt derived, driven by spring melt from the high-elevation Colorado Plateau winter snowpack. While winter precipitation provides only 30% of annual averages, it provides the majority of water for natural reservoirs [@sheppard2002].

### Data {#sec-data}

Daily observed streamflow data obtained from the United States Geological Survey (USGS) National Water Information System (NWIS) were used in this study. Streamgages were selected depending on criteria to ensure the applicability of each site. @odonnell2016 found that 8-10 years of calibration data were necessary to account for climate variability in paired watershed studies in the region. Following this criteria, a minimum of 10 years of record was chosen. Any years missing more than 30 days of streamflow were removed from the possible record period. This analysis focuses on natural, streamflow-influencing recharge, excluding streamgages affected by regulation or diversions. These gages were identified through annual reports on water data published by the USGS [@usgs2010]. Finally, streamgages on the Colorado River were excluded from the dataset on the basis that they represent managed flows shared by Colorado River Compact states. Following these parameters, 205 USGS streamgages with acceptable periods of record were used in this study (@fig-study-area). Periods of record ranged from 10 to 112 years, with a median of 28 years.

Watersheds across the United States are delineated by the USGS using a hydrologically-defined network. This system delineates the country using hierarchical hydrologic unit codes (HUCs), where each subsequent basin includes the digits of the enclosing basin. Here, 8-digit HUCs (HUC 8s) are used to divide Arizona into 84 sub-basins that are fully or partially in the state (@fig-study-area). These HUC 8 sub-basins are analogous to medium-sized river basins and are defined by surface water characteristics.

Annual precipitation and temperature data came from the PRISM climate group at Oregon State University at a resolution of 4 km (<https://prism.oregonstate.edu;> [@daly2008]). The PRISM dataset provides valuable insights into regional climate in ungauged regions and has been shown to perform well across the southwestern US [@buban_PRISM]**.** Instead of the water year, PRISM data uses a calendar-year format, which was adopted for consistency in the water balance. Although this may introduce challenges in the annual estimates due to inter-annual snow storage, the use of long-term annual averages is likely to reduce any potential errors [@reitz2017].

Annual evapotranspiration (ET) data came from TerraClimate, a 4-km grid climatalogical data set [@abatzoglou2018]. TerraClimate uses a Penman-Monteith approach to generate a reference potential evapotranspiration (PET). The PET values were calculated assuming a reference grass surface across the landscape with unlimited water. In the drylands of the southwestern US, PET typically exceeds precipitation annually [@zomer2022].

A 30-meter resolution Digital Elevation Model (DEM) of Arizona was used to derive key basin characteristics: basin area, average slope, and the proportion of each basin oriented toward north or south aspects. Various geospatial variables, such as aspect, were disaggregated then averaged to assess the areal percentage of each sub-variable within individual HUC 8 basins. By calculating these percentages, we aimed to get a more comprehensive understanding of landscape composition across space. Land cover from USGS-NLCD (National Land Cover Database), hydrologic soil group from SSURGO (Soil Survey Geographic Database), and underlying geology and karst from USGS were all similarly averaged across the basins. Aggregating variables to align with the HUC 8 boundaries allowed for more precise predictions of BFI by integrating spatial variations within each basin.

### Base-flow separation {#sec-bf_sep}

Directly measuring base flow and BFI presents unique challenges [@eckhardt2008]. The technique chosen to separate base flow has been shown to affect results, and the choice of base-flow separation method is subjective since ‘true’ BFI values are not known [@beck2013]. However, many methods have been developed to estimate these values. These methods include the use of tracers [@gonzales2009], graphical interpolation [@instituteofhydrology1980; @sloto1996], and digital filters [@arnold1995; @eckhardt2005; @lyne1979; @nathan1990]. These techniques have varying levels of applicability depending on the spatial scale, time span, and the scope of the study. Comparisons of various base-flow separation techniques have been made in [@eckhardt2005; @eckhardt2008; @nathan1990] , but this study does not explore the superiority of different methods.

Base flow was calculated using a single-parameter, recursive digital filter technique from [@nathan1990]. This base-flow separation technique is based on a recursive digital filter used in signal analysis that separates high-frequency signals (quickflow) from low-frequency signals (base flow) [@lyne1979]. @eckhardt2023 noted that recursive digital filters lack a physical basis, but as the method is easy to automate, objective, and repeatable, it is appropriate for a regional-scale study. The Lyne-Hollick filter has been used in [@ARNOLD200021; @santhi2008; @bloomfield2009; @singh2018]. It takes the form of

$$
b = \alpha b_{k-1} + \frac{1-\alpha}{2}(Q_k + Q_{k-1})
$$ {#eq-bf-separate}

where $b$ is base flow, $\alpha$ is the filter parameter, $Q$ is the total streamflow, and $k$ is the time step. A filter parameter $\alpha$ of 0.925 was used as in @nathan1990 and @fuka2014ecohydrology. The filter was run three times (forward, backward, forward) to attenuate the base-flow signal.

### Machine Learning

The implementation of machine learning models to predict hydrologic indices has been successful in past studies [@singh2018; @schmidt_ML_2020; @Rozos2021Machine]. In this work, we used the eXtreme Gradient Boosting (XGBoost) algorithm [@chen2016xgboost] to predict BFI at ungauged locations using catchment characteristics as predictors. The XGBoost algorithm is a decision tree-based ensemble algorithm, which can be adapted for either regression or classification problems. It has grown in popularity in recent years due to its rapid, efficient, and scalable nature, and has proven effective in similar environmental modeling tasks such as streamflow forecasting [@szczepanek2022daily**; @ni2020streamflow**] and classifying land use/land cover data [@Georganos2018Very].

XGBoost uses gradient boosting on decision tree algorithms to leverage the combination of multiple very low variance models into one overall prediction. Gradient boosting works sequentially, with the first tree being learned on the known target values, and each subsequent tree being learned on the errors of the tree prior. The algorithm then applies a weighting to each tree based on this boosting to determine how much the tree will contribute towards the final output. The final prediction is obtained by the input data traversing all $n$ weighted trees in the learned ensemble. The trained XGBoost model is used to predict BFI in ungauged catchments based on geospatial and hydroclimate predictor variables (@tbl-predictors). Certain features were subdivided according to their areal coverage of each basin (e.g. land cover was divided into 16 subdivisions).

|              | Variable                  | Source       | Geoprocessing          |
|------------------|-------------------|------------------|------------------|
| Hydroclimate | Precipitation\*           | PRISM        | Basin average          |
|              | Mean Temperature\*        | PRISM        | Basin average          |
|              | Mean Evapotranspiration\* | TerraClimate | Basin average          |
| Geospatial   | Elevation\*               | DEM          | Basin average          |
|              | Area                      | DEM          | Basin average          |
|              | Slope                     | DEM          | Basin average          |
|              | Aspect                    | DEM          | Percent areal coverage |
|              | Land Cover\*              | NLCD         | Percent areal coverage |
|              | Hydrologic Soil Group\*   | SSURGO       | Percent areal coverage |
|              | Geology                   | USGS         | Percent areal coverage |
|              | Karst                     | USGS         | Percent areal coverage |

: Basin-characteristic variables used as initial features in XGBoost model. Starred features are maintained in the final, dimensionality-reduced model. {#tbl-predictors}

Our training dataset consisted of 7724 observations of 45 variables. We first determined the optimal values for our hyperparameters using an exhaustive grid search and 5-fold cross-validation. The hyperparameters explored were: learning rate (referred to as $\eta$), minimum split loss ($\gamma$), maximum tree depth, minimum child weight, and number of trees. While these do not represent all possible hyperparameters of XGBoost, we are confident that the range of values explored represent sufficient variation to ensure that we have found a high-performing model. The optimal hyperparameters were found to be 700 trees, an $\eta$ of 0.05, a $\gamma$ of 0.075, maximum tree depth of 7, and a minimum child weight of 5. Using these values, the XGBoost model was then trained on the dataset using 10-fold cross-validation. $K$-fold cross-validation provides an unbiased estimate of a model's accuracy on unseen data, while also insuring against overfitting or underfitting. It does this by randomly splitting the data into $k$ folds of equal size, training the model on all-but-one fold, and testing on the one held-out fold (referred to as the validation set). This process is then repeated $k-1$ more times. We note that the training process starts fresh for each repeat of the process, meaning we are effectively training $k$ independent models, with each having the same hyperparameters but different training data. By averaging the performance of the model over $k$ folds we ensure robustness in our prediction accuracy. All of our models were optimized and assessed using root mean squared error as the performance metric.

#### Feature Selection

Machine learning models can face issues with overfitting when provided with a large set of predicting features. This can cause reductions in performance on unseen data and requires greater computational power and memory storage [@feat_selec2017]. Dimensionality reduction is a powerful tool in addressing these issues which can be divided into two broad categories: feature extraction and feature selection. Feature extraction involves projecting the higher dimensional dataset to a feature space with less dimensions, though this entails creating a new set of features that lose the physical meaning of the original feature space. On the other hand, feature selection maintains a subset of the original features, all of which retain their physical meaning providing better readability and interpretability. Here, a supervised features selection was used to reduce the number of predictors and, as a result, increased learning performance, decreased computational costs, and reduced overfitting.

An initial model was trained on the full feature set of 45 predictors (@tbl-predictors). We then used a feature importance method of feature selection to remove noise and less-important features. Feature importance scores are an approach where importance scores are assigned to individual features indicating how much the feature contributed (positively or negatively) to the model [@ML_interpretability2019]. Here, we use SHAP (SHapley Additive exPlanations) values as the method of assigning importance values to each feature [@SHAP_values]. The top-10 most-important features (@suppfig-shap_values) were subset and used in training a subsequent model which saw performance increases and reduced computing time. This subsequent model was used as the final model in this analysis.

### Statistical Analyses

Statisical analyses were performed on annual BFI and base flow values of instrumented streamgages. We examined temporal trends using the Mann-Kendall nonparametric trend test to determine the presence of monotonic trends [@kendall1970; @mann1945nonparametric]. The Mann-Kendall test was used to determine trends in BFI and trend significance of gauged reaches. This test is widely used in studies of this kind [@ficklin2016; @ayers2019; @woodhouse_udall22].

Trends in BFI and base flow were produced using a Mann-Kendall test for monotonic trends. This test expects that the data is non-parametric and that the observations are not autocorrelated over time. Using a Durbin-Watson test, observations at 4 streamgages show significant autocorrelation, though only one streamgage (09486500 - Santa Cruz River at Cortaro, AZ) indicated a significant trend in BFI. Due to this autocorrelation, the streamgage was removed from the trend analysis as the autocorrelation may inflate the variance of the Mann-Kendall statistic leading to an over/underestimate of the trend [@MK_Hamed1998].

Trends with a $\rho \le 0.05$ are considered significant.

## Results

### BFI of Gauged Catchments

![Mean annual BFI for the period of record from instrumented stream flow data.](images/BFI_Instrumented_20241203.png){#fig-instrumented-bfi}

The long-term average BFI for the 205 gauged reaches across Arizona is illustrated in @fig-instrumented-bfi . The highest BFI values (\>0.9) are found along the Grand Canyon in northwestern Arizona. The highly karstic geology of this region facilitates the rapid movement of subsurface flow to surface water and spring outlets [@chambless_deep-karst_2023]. Relatively high BFI values (\>0.8) are found at the spring-fed headwaters of the Verde River (Del Rio Spring) and the spring-fed headwaters of Fossil Creek. These results are consistent with interpolated BFI values reported by @wolock2003.

The stream reaches of the Little Colorado River Basin (northeastern Arizona) indicate consisently low BFI values (\< 0.2). This is likely due to low-yielding perched aquifers underlying the Defiance Uplift in northeastern Arizona, which are hydrologically connected to surface streams, while the high-yield, confined regional aquifer is much deeper [@blanchard02]. This region also typifies a pattern of seen along most major rivers in the study area: upstream reaches indicate higher BFI values while the downstream reaches display lower BFI values. This is assumed to be due to greater groundwater-surface water interaction at the headwaters of streams (likely due to spring outlets) and the dilution of base flow while moving down stream (**CITATION**). This pattern is seen along the Gila River, Verde River, and Little Colorado Rivers in Arizona.

#### Trends in BFI {#sec-bfi-trends}

Trends in BFI over the period of record for each streamgage in this analysis are illustrated in @fig-instrumented-trend and @tbl-trends .

![Trends in BFI over full period of record for instrumented sites used in this study. Red upward (blue downward) arrows indicate an increasing (decreasing) trend at a significance level of 5%. White circles represent sites with no statistically significant trends.](images/BFITrends_20241203.png){#fig-instrumented-trend}

| Classification Group               | n   | n_pos | n_neg | perc_pos | perc_neg |
|------------------------------------|-----|-------|-------|----------|----------|
| Precipitation - Monsoon Dominated  | 87  | 8     | 21    | 0.092    | 0.241    |
| Precipitation - Snowmelt Dominated | 118 | 9     | 12    | 0.076    | 0.102    |
| PhysRegion - Basin&Range           | 156 | 14    | 26    | 0.090    | 0.167    |
| PhysRegion - CO Plateau            | 49  | 3     | 7     | 0.061    | 0.143    |
| Climate - Warm-Wet                 | 31  | 2     | 6     | 0.065    | 0.194    |
| Climate - Warm-Dry                 | 55  | 6     | 11    | 0.109    | 0.200    |
| Climate - Cool-Wet                 | 74  | 4     | 9     | 0.054    | 0.122    |
| Climate - Cool-Dry                 | 45  | 5     | 7     | 0.111    | 0.156    |
| Slope - High                       | 102 | 10    | 12    | 0.098    | 0.118    |
| Slope - Low                        | 103 | 7     | 21    | 0.068    | 0.204    |

: Comparison of trends for BFI for all sites split by various classifications. Only sites with a significant ($\rho \le 0.05$) trend are included here as established by a Mann-Kendall test for monotonic trends across the full period of record. n is the number of sites, n_pos (n_neg) is the number of sites with positive (negative) trends, perc_pos (perc_neg) is the percentage of n with a positive (negative) trend. {#tbl-trends}

### BFI of Ungauged Catchments

#### Model Validation

Predicted values of BFI are plotted against observed values for the entire period of record of the instrumented dataset in @fig-actual_predicted . The agreement between 'out of bag' (blind cross-validation of the model as if each site were ungauged) predicted and observed values is acceptable (R^2^ = 0.764) and indicates that the model performs well across the full dataset. The overall Root Mean Square Error (RMSE) is 0.129 and the overall percent bias (pbias) is -5.6. Model performance metrics across various classifications are summarized in @tbl-performance . The various performance metrics indicate that this regional model performs well across different spatial and climate classifications. Negative pbias values across all classifications, along with the overall pbias, suggest that the model consistently under-predicts BFI. Categories with relatively poor performances of R^2^ and Nash-Sutcliffe Efficiency (NSE) also show relatively large biases.

![Linear relationship between observed BFI and predicted BFI. The solid line is the 1:1 line, the dashed line is regressed to the data.](images/actual-predicted.png){#fig-actual_predicted}

| Classification Group         | n    | R2    | MSE   | RMSE  | MAE   | NSE   | pbias |
|------------------------------|------|-------|-------|-------|-------|-------|-------|
| Climate - Monsoon Dominated  | 3039 | 0.633 | 0.016 | 0.126 | 0.074 | 0.619 | -13.7 |
| Climate - Snowmelt Dominated | 4685 | 0.733 | 0.015 | 0.121 | 0.087 | 0.725 | -3.5  |
| PhysRegion - Basin&Range     | 6147 | 0.733 | 0.016 | 0.127 | 0.084 | 0.724 | -6.3  |
| PhysRegion - CO Plateau      | 1577 | 0.846 | 0.011 | 0.104 | 0.073 | 0.843 | -3.8  |
| Climate - Warm-Wet           | 1506 | 0.693 | 0.014 | 0.117 | 0.077 | 0.685 | -8.2  |
| Climate - Warm-Dry           | 2351 | 0.693 | 0.022 | 0.147 | 0.092 | 0.675 | -11.9 |
| Climate - Cool-Wet           | 2350 | 0.738 | 0.011 | 0.106 | 0.078 | 0.736 | -1.7  |
| Climate - Cool-Dry           | 1517 | 0.831 | 0.012 | 0.111 | 0.078 | 0.827 | -4.3  |
| Slope - High                 | 3795 | 0.776 | 0.012 | 0.111 | 0.079 | 0.771 | -3.3  |
| Slope - Low                  | 3929 | 0.724 | 0.018 | 0.133 | 0.085 | 0.713 | -9.1  |

: Performance of model predictions for BFI for all sites split by various classifications. n is number of observatios, R^2^ is the coefficient of determination of a linear regression, MSE is mean-squared-error, RMSE is root-mean-squared-error, MAE is mean-absolute-error, NSE is Nash-Sucliffe efficiency, and pbias is percent bias. {#tbl-performance}

#### Predictor Importance {#sec-predictor-importance}

The predictors used to determine ungauged BFI were assessed to determine their importance. @suppfig-shap_values indicates the relative importance of predictors in the final XGBoost model along with the values of the . The most important feature in predicting long-term BFI is the elevation of the basin. While elevation has no physical impact on base flow characteristics, elevation has been shown to be one of the most important predictors in previous BFI studies [@singh2018; @beck2013]. The importance of catchment elevation is likely cumulative due to lower temperatures, orographic precipitation, and the persistence of snowpack at higher elevations. Seasonal snowpack duration has been shown to have a significant relationship to springflow, and thus groundwater recharge, in the region [@donovan_karst_2022]. This relationship is evident in further hydroclimate features. Higher temperatures tend to have a negative influence on BFI, while precipitation shows a mixed influence with some higher values negatively influencing BFI. This is likely due to a greater proportion of this precipitation leading to runoff, thus resulting in lower BFI values.

The features describing land cover and land use indicate that a higher percentages of evergreen forest have a positive influence on predicted BFI. On the other hand, higher percentages shrub-land and developed land have negative influences on BFI. Hydrologic soil types also show strong trends in their influence on BFI. Soil-type C indicates soils with moderately high runoff potential when wetted, typically having 20-40% clay. Soil-type A has low runoff potential, facilitating rapid transmission of water through the soil profile [@nrcs2009]. With this being the case, higher percentages of soil-type C tend to have a negative influence on BFI while soil-type A percentages show a mixed influence. This mixed influence is likely due to SHAP values reflecting interactions between features in the model as opposed to direct relationships between feature and prediction. Regions where soil-type A dominate may be found along steep slopes or low vegetation which dominate the model prediction. Furthermore, the relationship between soil-type A and BFI may be non-linear and more complex in regional base-flow dynamics.

#### Predicted Long-term Mean BFI

The long-term BFI (1991-2020) for all stream reaches in Arizona of Strahler stream order 3 and greater are shown in @fig-bfi-streams and the regionalized (HUC-8) long-term BFI is shown in @fig-bfi-huc .

![Predicted long-term BFI values for 8-digit HUC (1991-2020)](images/BFI_HUC8_20241203.png){#fig-bfi-huc}

![Predicted long-term BFI values for all stream reaches of Strahler stream order 3 and greater (1991-2020)](images/BFI_AllStream.png){#fig-bfi-streams}

## Conclusion {#sec-conclusion}

-   This regional model, trained on a relatively sparse dataset over a region spanning 295,253 km^2^, is shown to predict BFI well spatially and across climate classifications.

-

## References {.unnumbered}

## Supporting Information

![SHAP value plot of features used in final model. LC indicates land cover, and soil types are defined by hydrologic soil group.](images/supplemental/shap_summary.png){#suppfig-shap_values}