### 2025-01-20 Supp_Info.qmd
supp info

---
title: "Supplemental Info"
editor: visual
---

![Histogram of BFI trends grouped by long-term instrumented BFI. Significant trends (\>0.05) are indicated with bolded borders. More significant negative trends are apparent in low-BFI gauges. Significant positive trends are equally distributed among BFIs.](images/supplemental/bfi-trend_histogram.png)

![Quadrant plot of Kendall's Tau trends in base flow vs. BFI. Significant trends in base flow and BFI indicate that variations in base flow volumes are coincident with trends in BFI. This indicates that trends in BFI typically follow trends in groundwater discharge (base flow) and that trends in BFI aren't typically due to an overall decrease/increase in streamflow. One site indicates a decrease in base flow volumes and an increase in BFI, suggesting that decreases in surface water inputs is of a greater magnitude than the increase of base flow.](images/supplemental/baseflow-bfi_quadrants.png)

![Plot of residuals of predicted BFI vs observed BFI.](images/supplemental/act-pred_residuals.png)